Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-28 Response.tex

                    % draw the subsegments
                    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
                    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
                    % dotted lines here
                    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);

                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
                \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Schematic diagram of the multiple shooting method:
        The complete trajectory is split into a number of sub-segments, and additional interior constraints are included to ensure state and costate continuity.
    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
\end{figure}

\begin{figure}[H]
   \centering
   \includegraphics[width=0.2\textwidth]{moon_orbit} % requires the graphicx package
   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstrating the initial and target orbit.
   Without the low-thrust propulsion, the spacecraft is constrained to the initial periodic orbit. 
   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
\end{figure}

\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{manifold_trajectory} 
                \caption{Position space view of invariant manifold} 
        \end{subfigure}~ %
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{manifold_poincare} 
                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} 
        \end{subfigure} 
        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
\end{figure}
\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{reach_trajectory} 
                \caption{Position space view of reachability set}  
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{poincare_compare} 
                \caption{\Poincare section view of reachability set}  
        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

### 2017-07-30 Response.tex

\item 
    \begin{itshape}
The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
\end{itshape}

Thank you for the comment, however it seems to be rare for the research community to give detailed description of the numerical optimization method.
Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software. 
There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
However, our manuscript is primarily focused on demonstrating the feasibility of using reachability sets to design transfer trajectories.
As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.
However, we have added some rough detail on the computation time and computer hardware used in the numerical optimization.
This additional detail is copied below.

\begin{correction}
Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
\end{correction}

### 2017-07-30 manuscript.tex
Remove gradient comment and update timing of computation

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
Each solution on the reachability set is computed in approximately \SI{30}{\second} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.

>>>>

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
We discretely vary the angle over the range \( \ang{0} \leq \theta_d < \ang{360} \) to approximate the reachability set of the spacecraft.
Choosing a new angle \( \theta_d \) corresponds to a different direction as well as a new optimal control problem which is again solved using the multiple shooting approach laid out previously.
Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
Each solution on the reachability set is computed in approximately \SI{2}{\minute} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:invariant_manifold_transfer}.
